---
title: "Lab 6 Comparing two means"
author: "Phillip Hagen"
output: github_document
---
```{r}
library(tidyverse)
```

Researchers studying the number of electric fish species living in various parts of the Amazon basin were interested in whether the presence of tributaries affected the local number of electric fish species in the main rivers (Fernandes et al. 2004).

They counted the number of electric fish species above and below the entrance point of a major tributary at 12 different river locations.

The data is provided in your GitHub repository.

For each question below, write a sentence answering the question and show the code you used to come up with the answer, if applicable.

```{r}
fish <- read_csv("chap12q19ElectricFish.csv")

fish_long <- 
  pivot_longer(fish, speciesUpstream:speciesDownstream,
               names_to = "location",
               values_to = "species") %>% 
  mutate(location = str_remove(location, c("species"))) %>% 
  print()

```

## Question A

> What is the mean different in the number of species between areas upstream and downstream of a tributary? What is the 95% confidence interval of this mean difference. Show your code and write a sentence giving your answer

ANSWER

The mean difference is 1.83334, and the 95% confidence interval of the mean difference is -4.587031 to 8.253697. 
```{r}
fish_long_summary <-
  fish_long %>% 
  group_by(location) %>% 
  summarize(
    n = n(),
    mean = mean(species),
    sd = sd(species),
    sem = sd/sqrt(n),
    upper = mean + 1.96 * sem,
    lower = mean - 1.96 * sem
  ) %>% 
  print()

t.test(formula = species ~ location, data = fish_long)

16.41667-14.58333
```

## Question B

> Test the hypothesis that the tributaries have no effect on the number of species of electric fish.

ANSWER

The p-value is 0.5596, therefore we would fail to reject the null hypothesis. 

We can see from the graph below that there is quite a bit of overlap between downstream and upstream. This means there is no normal distribution. 
```{r}
fish_long %>% 
  ggplot(aes(x = location, y = species)) +
  geom_jitter(aes(color = location), 
              shape = 16, size = 3, 
              alpha = 0.3, width = 0.4) +
  geom_errorbar(aes(y = mean, ymax = upper, ymin = lower), 
                data = fish_long_summary, 
                width = .1, size = .8) +
  geom_point(aes(y = mean), 
             data = fish_long_summary, 
             size = 3) +
  scale_color_manual(values = c("darkorange","cyan4")) +
  theme_minimal() +
  guides(color = "none")
```

## Question C

> State the assumptions that you had to make to complete parts (A) and (B). Create a graph to assess whether one of those assumptions was met.

ANSWER

To complete parts A and B we needed to assume that there would be a normal distribution between downstream and upstream. 

From the graph we can see the distribution does not appear to be normal which means our assumptions were not met. 
```{r}
fish_long %>% 
  ggplot(aes(x = species)) +
  geom_histogram(
    aes(fill = location), 
    bins = 15, 
    alpha = 0.5, 
    position = "identity"
  ) +
  scale_fill_manual(values = c("darkorange","cyan4")) +
  theme_minimal()
```

## ANOVA

Fiddler crabs are so called because males have a greatly enlarged "major" claw, which is used to attract females and to defend a burrow.

Darnell and Munguia (2011) recently suggested that this appendage might also act as a heat sink, keeping males cooler while out of the burrow on hot days.

To test this, they placed four groups of crabs into separate plastic cups and supplied a source of radiant heat (60-watt light bulb) from above. The four groups were intact male crabs, male crabs with the major claw removed; male crabs with the other (minor) claw removed (control); and intact female fiddler crabs.

They measured the body temperature of crabs every 10 minutes for 1.5 hours. These measurements were used to calculate a rate of heat gain for every individual crab in degrees C/log minute. Rates of heat gain for all crabs are provided in the accompanying data file.

```{r}
crab <- read_csv("chap15q27FiddlerCrabFans.csv")
```


### Question D

> Graph the distribution of body temperatures for each crab type:

```{r}
crab %>% 
  ggplot(aes(x = bodyTemperature)) +
  geom_histogram(
    aes(fill = crabType),
    bins = 12, 
    alpha = 0.5, 
    position = "identity", 
    na.rm = TRUE
  ) +
  scale_fill_manual(values = c("darkorange","cyan4", "red", "green")) +
  theme_minimal()
```

### Question E

> Does body temperature varies among crab types? State the null and alternative hypothesis, conduct and ANOVA, and interpret the results.

The null hypothesis was that body temperature between crabs would not vary. The alternative was that body temperature would be different between one crab type. 
We reject the null hypothesis and accept the alternative. This is because there was a statistical significance between body temperature among the crab types. We determined this from F being greater than one and the p-value being less than 0.05. 
```{r}
aov_crabby <-
  aov(bodyTemperature ~ crabType, data = crab)
aov_crabby

summary(aov_crabby)
```